## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}
a <-10

if(a<=5){
  print (a)
} else {
  print('it more than 5')
}
```

We can also add else if statements in the middle if we need more options.

```{r}
a <- 15

if (a<=5){
  print('small')
} else if (a>5 & a <=10){
  print('medium')
} else if (a>10 | a ==7){
  print('large')
}else{
  print('huge')
}
```

Neat! We can also nest conditionals just like we do in Python.

```{r}
a <- 3
if(a/2 != 0){
  if (a ==5){
   print('a is 5') 
  }
}

```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}
for(value in 1:5){
  print (paste(value, 'is greater than', value -1 ))
}
```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}
a_vector <- c(1,1,1,1,1,1,1,1)

for (i in c(1,3,6)){
  a_vector[i] <- 6
}
a_vector
```


We can loop over more than just lists:

```{r}
mat <- matrix(1, nrow =2, ncol = 2)
mat

for(row in 1:2){
  for(col in 1:2){
    mat[row,col] <- 0
    print(mat)
  }
}
```

Like Python, we have while loops in R:

```{r}
b <- 0 
while (b<10) {
  b <- b+2
  print(b)
}
```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}
test_temp<- c(16.55,26.94,32.59)

c_to_f <- function(dat){
  temp_fahr <- dat*(9/5) +32
  
  return(temp_fahr)
}


```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}
data_fahr <- c_to_f(test_temp)
data_fahr
```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}
test_mat <- matrix(c(1,1,1,2,2,2), nrow <-2)

apply(test_mat,1,sum)
```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}
test_temp_mat<- matrix(c(16.55,26.94,32.59))

apply(test_temp_mat ,1,c_to_f)
```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
#In this simple assignment you are given a number and have to make it negative. But maybe the number is already negative?
    
    make_negative <- function(num){
  if (num>0){
    print(num * -1)
  } else if (num<0){
    print (num * 1)
    }else{
      print (num)
    }
  }
    
 
    
#Make a function that will return a greeting statement that uses an input; your program should return, "Hello, <name> how are you doing today?".
    greet <- function(name) {
    return (sprintf("Hello, %s how are you doing today?", name))
    
}
  
    
#Implement a function which convert the given boolean value into its string representation.
     boolean_to_string <- function(b){
   return (as.character(b))
     }
    
     
#Write function RemoveExclamationMarks which removes all exclamation marks from a given string.
     remove_exclamation_marks <- function(s){
  return(gsub('[!]','', s))
     }
     
#Timmy & Sarah think they are in love, but around where they live, they will only know once they pick a flower each. If one of the flowers has an even number of petals and the other has an odd number of petals it means they are in love.

#Write a function that will take the number of petals of each flower and return true if they are in love and false if they aren't.
     
      lovefunc <- function(flower1, flower2){
    return ((flower1 + flower2) %% 2 != 0)
  }
    
    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
    number<- function(num){
    if(num %% 3 == 0 & num %% 5 == 0){
      print('fizzbuzz')
    } else if(num %% 3 == 0){
      print('fizz')
      }else if(num %% 5 == 0){
      print('buzz')
    }
    }
    
    number(30)

    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.
    
    mariostairs <- function(n){
      for (i in 1:n) {
          cat(strrep('x',i),sep='\n')
        }
      }
    
     
    
      
      
    
          
     mariostairs(4)
     
     
     
     
    ```
